ProbabilityBoundsAnalysis.jl:\\ an arithmetic of probabilities and sets of probabilities

### 2021-09-27 paper/paper.tex
finished bivariate p-boxes

\begin{equation*}
  [\underline{F}^{-1}(\alpha), \overline{F}^{-1}(\alpha)]
\end{equation*}

@@

where the max operator is required when $\underline{F}(b) < \overline{F}(a)$. Note the same can be achieved by using the standard formula for finding the measure from the cdf, $\mathbb{P}(U) = F(b)- F(a)$, and using interval arithmetic.

@@

These p-boxes are constructed using the Chebyshev (mean and variance), Markov (mean and one bound) and Cantelli (mean and range) inequalities from probability theory. These inequalities allow for the cdf to be bounded from various range and moment information. For example, the Chebyshev inequality can be used to find p-box bounds from the random variables mean and variance:

@@

Samples and the probability measure can also be found from a bivariate p-box. Like how a sample of a univariate p-box is an interval, a sample of a bivariate p-box will be an interval box, a two dimensional rectangular set.

A minimal structure and cdf of a bivariate p-box is

@@

  function cdf(F :: BivPbox, x, y)

@@

    return IntervalBox.(X_samps, Y_samps)
  end
\end{lstlisting}

>>>>

\begin{equation*}
  [\overline{F}^{-1}(\alpha), \underline{F}^{-1}(\alpha)]
\end{equation*}

@@

where the max operator is required when $\underline{F}(b) < \overline{F}(a)$. Note the same can nearly be achieved by using the precise formula for the probability measure, $\mathbb{P}(U) = F(b)- F(a)$, and using interval arithmetic.

@@

These p-boxes are constructed using the Chebyshev (mean and variance) \cite{chebyshev1874valeurs}, Markov (mean and one bound) \cite{markoff1900question} and Cantelli (mean and range) inequalities from probability theory. These inequalities allow for the cdf to be bounded from various range and moment information. For example, the Chebyshev inequality can be used to find p-box bounds from the random variables mean and variance:

@@

Samples and the probability measure can also be found from a bivariate p-box. Like how a sample of a univariate p-box is an interval, a sample of a bivariate p-box will be an interval box, i.e. a two dimensional rectangular set. Multiple samples will therefore be a population of interval boxes, which together follow the dependency encoded in $C$. The bottom of Figure \ref{fig:figure7} shows 100 samples of a bivariate p-box, whose marginals are $X \sim \text{Beta}([4,8], 3)$ and $Y \sim \text{N}([0,2], 2)$, and with the Gaussian copula (multivariate normal's copula) with correlation coefficient $\rho_{XY} = -0.8$: $C = \Phi_{-0.8}$. The top of Figure \ref{fig:figure7} shows the cdf bounds. The probability measure from a bivariate p-box in the rectangular set $ U = [x_{1}, x_{2}] \times [y_{1}, y_{2}]$ is\footnote{Which is nearly a straightforward application of interval arithmetic to the precise calculation $F(x_{2}, y_{2}) - F(x_{1}, y_{2}) - F(x_{2}, y_{1}) + F(x_{1}, y_{1})$}:

\begin{align*}
  \underline{\mathbb{P}}(U) &= \text{max}(0, \underline{F}(x_{2}, y_{2}) - \overline{F}(x_{1}, y_{2}) - \overline{F}(x_{2}, y_{1}) + \underline{F}(x_{1}, y_{1})) ,\\ 
  \overline{\mathbb{P}}(U)  &= \text{min}(1, \overline{F}(x_{2}, y_{2}) - \underline{F}(x_{1}, y_{2}) - \underline{F}(x_{2}, y_{1}) + \overline{F}(x_{1}, y_{1}) ).
\end{align*}

A minimal structure, cdf, random sample and probability measure of a bivariate p-box is:

@@

  function cdf(F :: BivPbox, x :: Real, y :: Real)

@@

    return IntervalBox.(X_samps, Y_samps)
  end

  function mass(F :: BivPbox, box :: IntervalBox)

    x = box.v[1]; y = box.v[2];  # gets intervals

    F_1 = cdf(F, x.lo, y.lo)
    F_2 = cdf(F, x.hi, y.lo)
    F_3 = cdf(F, x.lo, y.hi)
    F_4 = cdf(F, x.hi, y.hi)

    # using intervals
    mass = F_4 - F_3 - F_2 + F_1;

    return max(0, min(1, mass))
  end
\end{lstlisting}

  \includegraphics[width=0.5\textwidth]{../examples/JuliaCon/fig7/biv_cdf.pdf}

  \includegraphics[width=0.5\textwidth]{../examples/JuliaCon/fig7/samples.pdf}

  \caption{Cdf bounds of a bivariate p-box (top) with $X \sim \text{Beta}([4,8], 3)$, $Y \sim \text{N}([0,2], 2)$, and $\rho_{XY} = -0.8$, and 100 samples (bottom)}
  \label{fig:figure7}
